The Oumupo, between music and math

### 2017-05-03 article.tex
finishes martin's section

We all know the small game of folding a slip of paper $n$ times in the same
direction and then unfolding the paper to see the final form.  The
complexity of the exercise increases according to the number of folds.  This
Dragon Curve possesses several interesting properties (fractal,
self-similar, symmetric but non-periodic, etc.).  The curve can also be read
as a series of melodic movements
\begin{enumerate}
\item Choose a scale, with a preference for non-standard scales such as C
D$\sharp$ E F G A$\flat$ B.
\item Fold a slip of paper twice in the same direction, then unfold it and
lie it down flat.
\item Play the first note of the scale, and for each ``bump'' in the paper,
go up one note.  For each dip in the paper, go down by one note (always
following the chosen scale).
\item Take the slip of paper, add an extra fold, and play the new melody
obtained by this change.
\end{enumerate}
You will notice that the melody becomes more and more complex, but also more
and more maze-like and mysterious.  The breathtaking and hypnotic beauty of
the Dragon Curve allows one to ``hear'' mathematical constructs in exciting
new ways.
\section*{The conquest Niagra: a radical approach to sound compression}

\end{document}

>>>>

We all know the small game of folding a slip of paper $n$ times in the same
direction and then unfolding the paper to see the final form.  The
complexity of the exercise increases according to the number of folds.  This
Dragon Curve possesses several interesting properties (fractal,
self-similar, symmetric but non-periodic, etc.).  The curve can also be read
as a series of melodic movements
\begin{enumerate}
\item Choose a scale, with a preference for non-standard scales such as C
D$\sharp$ E F G A$\flat$ B.
\item Fold a slip of paper twice in the same direction, then unfold it and
lie it down flat.
\item Play the first note of the scale, and for each ``bump'' in the paper,
go up one note.  For each dip in the paper, go down by one note (always
following the chosen scale).
\item Take the slip of paper, add an extra fold, and play the new melody
obtained by this change.
\end{enumerate}
You will notice that the melody becomes more and more complex, but also more
and more maze-like and mysterious.  The breathtaking and hypnotic beauty of
the Dragon Curve allows one to ``hear'' mathematical constructs in exciting
new ways.
\section*{The conquest of Niagra: a radical approach to sound compression}
The entire history of musical comression can be summed up in a single
phrase: putting as much information as possible into the smallest space
possible.  But why on Earth would anyone want to fill a hard drive with ten
years of music?  No one would ever spend ten years of her life actually
listening to the stuff (how outrageous!).  It is time to find innovative
compression solutions that actually correspond to people's needs.
The \textbf{acceleration} of recorded music is a compression method as old
as the record itself.  In playing a disk 33 rotations-per-minute record
at a speed of 45 RPM, one achieves a time reduction of 36\% (an acceleration
of 136\% is possible at 78 RPM).  Of course, analog time-compression in
music results pushes the original frequencies higher, meaning that only
certain pieces lend themselves to this treatment -- for example, Erik
Satie's 3rd \emph{Gymnop\'edie}, entitled ``slow and grave.''

The more difficult method of \textbf{superposition} requires searching
similar passages in one or several works and playing them simultaneously. 
An interesting research subject is the \emph{anatole}, the chord progression
that is the bread and butter of jazz standards and popular music.  By
superposing all of the works constructed using this sequence, one can gain a
lot of time!

Last but not least, \textbf{random superposition} allows extremely efficient
musical gains but significantly lowers the music's signal-to-noise ration.
For example, H.I.F von Biber's \emph{La Battalia à 10} superposes eight
popular melodies in different tonalities and tempos to evoke the theme of
drunken soldiers.  The 8/1 compression allows for an 87.5\% reduction in
time.  In a similar vein, the piece \emph{Folk Music} by Zygmunt Krause
proposes a superposition of twenty folk melodies -- a whopping 95\%! Pushed
to its extreme, this logic allows one to listen to all possible music at the
same time.  Also known as ``white noise,'' this is compression at is limit,
as the quantity of musical information tends towards zero.  In other words,
compression moves music towards the sound of Niagra Falls.
\end{document}
